Phylodynamic inference of contact network parameters through approximate Bayesian computation

### 2016-06-27 doc/thesis/methods_results.Rnw
Paper and defence presentation updates

\subsection{Using synthetic data to investigate identifiability and sources of
estimation error and bias}

@@

The accuracy of classifiers for $I$ \del{varied significantly
with}\add{appeared to be strongly influenced by} the number of tips in the
tree. For 100-tip trees, the average $R^2$ values were
  \Sexpr{kernel.I[ntip == 100, round(mean(rsquared), 2)]},
  \Sexpr{nltt.I[ntip == 100, round(mean(rsquared), 2)]}, and
  \Sexpr{sackin.I[ntip == 100, round(mean(rsquared), 2)]}
for the tree kernel, \gls{nltt}, and Sackin's index respectively. For 500-tip
trees, the values increased to
  \Sexpr{kernel.I[ntip == 500, round(mean(rsquared), 2)]},
  \Sexpr{nltt.I[ntip == 500, round(mean(rsquared), 2)]}, and
  \Sexpr{sackin.I[ntip == 500, round(mean(rsquared), 2)]}.
Finally, the performance of classifiers for $N$ depended heavily on the
epidemic scenario. The $R^2$ of the \gls{kSVR} classifier ranged from
  \Sexpr{kernel.N[,round(min(rsquared), 2)]}
for the smallest epidemic and smallest sample size, to
  \Sexpr{kernel.N[,round(max(rsquared), 2)]}
for the largest. Likewise, $R^2$ for the \gls{nltt}-based SVR ranged from 
  \Sexpr{nltt.N[,round(min(rsquared), 2)]}
to
  \Sexpr{nltt.N[,round(max(rsquared), 2)]}.
Sackin's index did not accurately classify $N$ in any scenario, with an average
$R^2$ of
  \Sexpr{sackin.N[,round(mean(rsquared), 2)]}
and little variation between scenarios.

@@

\begin{table*}[ht]
    \centering
    \input{\tablepath/abc-hpd.tex}
    \caption[
        Average posterior mean point estimates and 95\% \glspl{HDI} interval
        widths for \gls{BA} model parameter estimates obtained with
        \software{netabc} on simulated data.
    ]{
        Average posterior mean point estimates and 95\% \glspl{HDI} interval
        widths for \gls{BA} model parameter estimates obtained with
        \software{netabc} on simulated data. Three transmission trees were
        simulated under each combination of the listed parameter values, and
        the parameters were estimated with \gls{ABC} without training.
    }
    \label{tab:abchpd}
\end{table*}

\add{Across all simulations, the median [\gls{IQR}] absolute errors of the parameter 
estimates obtained with \software{netabc} were
    \Sexpr{d[,round(median(alpha_error), 2)]} 
    [\Sexpr{d[,round(quantile(alpha_error, 0.25), 2)]} - 
    \Sexpr{d[,round(quantile(alpha_error, 0.75), 2)]}]
for \gls{alpha},
    \Sexpr{d[,round(median(I_error))]} 
    [\Sexpr{d[,round(quantile(I_error, 0.25))]} - 
    \Sexpr{d[,round(quantile(I_error, 0.75))]}]
for \gls{I},
    \Sexpr{d[,round(median(m_error), 2)]} 
    [\Sexpr{d[,round(quantile(m_error, 0.25), 2)]} - 
    \Sexpr{d[,round(quantile(m_error, 0.75), 2)]}]
for \gls{m}, and
    \Sexpr{d[,round(median(N_error))]} 
    [\Sexpr{d[,round(quantile(N_error, 0.25))]} - 
    \Sexpr{d[,round(quantile(N_error, 0.75))]}]
for \gls{N}. These errors comprised, respectively,
    \Sexpr{d[,round(median(alpha_error) * 100 / 2)]}\%,
    \Sexpr{d[,round(median(I_error) * 100 / 4500)]}\%,
    \Sexpr{d[,round(median(m_error) * 100 / 6)]}\%, and
    \Sexpr{d[,round(median(N_error) * 100 / 14500)]}\%
of the regions of nonzero prior density. For \gls{I} and \gls{N}, relative
errors were
    \Sexpr{d[,round(median(I_error * 100 / true_I))]}\%
    [\Sexpr{d[,round(quantile(I_error * 100 / true_I, 0.25))]} - 
    \Sexpr{d[,round(quantile(I_error * 100 / true_I, 0.75))]}\%]
and
    \Sexpr{d[,round(median(N_error * 100 / true_N))]}\%
    [\Sexpr{d[,round(quantile(N_error * 100 / true_N, 0.25))]} - 
    \Sexpr{d[,round(quantile(N_error * 100 / true_N, 0.75))]}\%].
Average \gls{HPD} interval widths were
    \Sexpr{d[,round(mean(alpha.upper.95 - alpha.lower.95), 2)]},
    \Sexpr{d[,round(mean(I.upper.95 - I.lower.95))]},
    \Sexpr{d[,round(mean(m.upper.95 - m.lower.95), 2)]}, and
    \Sexpr{d[,format(round(mean(N.upper.95 - N.lower.95)), digits=5)]},
representing 
    \Sexpr{d[,round(mean(alpha.upper.95 - alpha.lower.95) * 100 / 2)]}\%,
    \Sexpr{d[,round(mean(I.upper.95 - I.lower.95) * 100 / 4500)]}\%,
    \Sexpr{d[,round(mean(m.upper.95 - m.lower.95) * 100 / 6)]}\%, and
    \Sexpr{d[,round(mean(N.upper.95 - N.lower.95) * 100 / 14500)]}\%
of the nonzero prior density regions. Point estimates of \gls{I} were upwardly
biased: \gls{I} was overestimated in
    \Sexpr{d[,sum(I > true_I)]}
out of 
    \Sexpr{nrow(d)}
simulations
    (\Sexpr{round(d[,sum(I > true_I)] * 100 / nrow(d))}\%).
The estimates for \gls{m} and \gls{N} were similar across all simulations
(median [IQR] point estimates
    \Sexpr{d[,round(median(m), 2)]}
    [\Sexpr{d[,round(quantile(m, 0.25), 2)]} - 
    \Sexpr{d[,round(quantile(m, 0.75), 2)]}]
and 
    \Sexpr{d[,round(median(N))]}
    [\Sexpr{d[,round(quantile(N, 0.25))]} - 
    \Sexpr{d[,round(quantile(N, 0.75))]}])
regardless of the true values of any of the \gls{BA} parameters.}

\add{To analyze the effects of the true parameter values on the
accuracy our estimates of \gls{alpha} and \gls{I}, we fitted one \glspl{GLM}
for each of these two parameters, with error rate as the dependent variable and
the true parameter values as independent variables. Since the estimates of
\gls{m} and \gls{N} were roughly equal across all simulations
(\cref{fig:abcptm2,fig:abcptm3,fig:abcptm4}), \glspl{GLM} were not fitted for
these parameters. The estimated coefficients are shown in
\cref{tab:glmalpha,tab:glmI}. For the parameters \gls{alpha}, \gls{I}, and
\gls{N}, the \gls{GLM} were fitted using the inverse link function. That is, if
$p$ is the true value of the parameter and $\hat{p}$ is a random variable
representing our estimate of the parameter, the \gls{GLM} posits a relationship
of the form}
\[
    \madd{\E(|p-\hat{p}|) = (\beta_0 + \beta_\alpha \alpha + \beta_I I + \beta_m m)^{-1}},
\]
\add{where the $\beta$'s are coefficients to be fitted. If the true value
\gls{alpha}, say, is increased by one, the \emph{inverse} of the expected
absolute error will increase by $\beta_\alpha$. If $\beta_\alpha$ is positive,
it means that the absolute error decreases as the true value of \gls{alpha}
increases.}

@@

<<glm_tables, echo=FALSE, warning=FALSE, error=FALSE, results="asis">>=
    source("global.R")
    format.glm <- function (df) {
        df.copy <- copy(df)
        df.copy[Parameter == "alpha", Parameter := "$\\alpha$"]
        df.copy[,p := paste0("$", sapply(p, pp, eq=FALSE), "$")]
        df.copy[,Estimate := paste0("$", sapply(Estimate, pp, eq=FALSE, thres=NA), "$")]
        df.copy[,`Standard error` := paste0("$", sapply(`Standard error`, pp, eq=FALSE, thres=NA), "$")]
        setnames(df.copy, "p", "$p$-value")
        df.copy
    }

@@

<<supp_labels, echo=FALSE, results="asis">>=
    #N <- 5000
    #replicate <- 0
    #lab <- NULL
    #for (m in c(2, 3, 4)) {
    #for (I in c(1000, 2000)) {
    #for (alpha in c(0, 0.5, 1, 1.5)) {
    #    lab <- c(lab, sprintf("fig:%.1f-%d-%d-%d-%d", alpha, I, m, N, replicate))
    #}
    #}
    #}
    #lab <- paste(lab, collapse = ",")
    worst.alpha <- d[,mean(alpha_error), by=true_alpha][,true_alpha[which.max(V1)]]
    best.alpha <- d[,mean(alpha_error), by=true_alpha][,true_alpha[which.min(V1)]]
    worst.I <- d[,mean(I_error), by=true_I][,true_I[which.max(V1)]]
    best.I <- d[,mean(I_error), by=true_I][,true_I[which.min(V1)]]
    worst.m <- d[,mean(m_error), by=true_m][,true_m[which.max(V1)]]
    best.m <- d[,mean(m_error), by=true_m][,true_m[which.min(V1)]]

@@

The dispersion of the \gls{ABC} approximation to the posterior also varied
between the parameters (\cref{tab:abchpd}). \gls{HPD} intervals around
\gls{alpha} and \gls{I} were often narrow relative to the region of nonzero
prior density, whereas the intervals for \gls{m} and \gls{N} were more widely
dispersed. \Cref{fig:abclow1,fig:abclow2} \del{shows the distributions for one
simulation.} \add{show one- and two-dimensional marginal distributions for a
simulation with relatively low error. Each parameter was chosen based on its
mean error rate across all simulations. The parameters for this simulation were 
    $\gls{alpha} = \Sexpr{best[,true_alpha]}$,
    $\gls{I} = \Sexpr{best[,true_I]}$,
    $\gls{m} = \Sexpr{best[,true_m]}$, and
    $\gls{N} = \Sexpr{best[,true_N]}$.
\Cref{fig:abchigh1,fig:abchigh2} show the equivalent marginals for a different
simulation with relatively high error rates for each individual parameter. The
parameters were
    $\gls{alpha} = \Sexpr{worst[,true_alpha]}$,
    $\gls{I} = \Sexpr{worst[,true_I]}$,
    $\gls{m} = \Sexpr{worst[,true_m]}$, and
    $\gls{N} = \Sexpr{worst[,true_N]}$.
The two-dimensional marginals some dependence between pairs of parameters,
particularly \gls{I} and \gls{N} which show a diagonally shaped region of high
posterior density.}

@@

\del{We applied \software{netabc} to five published HIV datasets (\cref{tab:data}).
and found substantial heterogeneity among the parameter
estimates.} \add{Posterior mean point estimates and 50\% and 95\% \glspl{HPD}
for each parameter are shown in \cref{fig:abchpd}.} \del{and marginal posterior
distributions in \cref{fig:cuevas,fig:li,fig:niculescu,fig:novitsky,fig:wang}}.
\add{\Cref{fig:abchpdm2} shows point estimates and \glspl{HPD} obtained when
the value $\gls{m} = 1$ was disallowed by the prior. Since the results
indicated that $\gls{m} = 1$ was the most credible value for several datasets,
all results discussed henceforth are for the prior $\gls{m} \sim
\DiscreteUniform(1, 5)$ unless otherwise stated.}

\begin{figure}[ht]
    \centering
    \includegraphics{realdata-hpd-bc}
    \caption[
        Posterior means and 50\%/95\% \gls{HPD} intervals for parameters of the
        \gls{BA} network model, fitted to eleven
        \gls{HIV} datasets with \software{netabc}.
    ]{
        Posterior means (points), 50\% \gls{HPD} intervals (notches), and 95\%
        (lines) \gls{HPD} intervals for parameters of the \gls{BA} network
        model, fitted to eleven \gls{HIV} datasets with \software{netabc}.
        Legend labels indicate risk group and country of origin. Abbreviations:
        \acrshort{IDU}, \acrlong{IDU}; \acrshort{MSM}, \acrlong{MSM}; HET,
        heterosexual. Note that posterior means can fall outside of \glspl{HPD} 
        if the distribution is diffuse.
    }
    \label{fig:abchpd}
\end{figure}

@@

\add{Posterior mean point estimates for the \acrlong{PA} power \gls{alpha}
were all sub-linear, ranging from 
    \Sexpr{pmean1["zetterberg2004", round(alpha, 2)]}
for the \gls{IDU}/Estonia data to
    \Sexpr{pmean1["cuevas2009", round(alpha, 2)]}
for the mixed/Spain data. When aggregated by risk group, the average estimates
were
    \Sexpr{pmean1[c("zetterberg2004", "niculescu2015", "bctree"), round(mean(alpha), 2)]}
for \gls{IDU},
    \Sexpr{pmean1[c("novitsky2014", "mccormack2002", "grabowski2014"), round(mean(alpha), 2)]}
for primarily heterosexual risk, and
    \Sexpr{pmean1[c("wang2015", "kao2011", "little2014", "li2015"), round(mean(alpha), 2)]}
for \gls{MSM}. These values were obtained with \textit{gag} for the datasets
where both \textit{gag} and \textit{env} were sequenced, but the estimates for
\gls{alpha} did not change appreciably between the two genes (\cref{fig:genes}).
There was a large amount of uncertainty associated with these estimates. 95\%
\gls{HPD} were very wide for most datasets, often encompassing almost the
entire range from 0 to 1 (\cref{fig:abchpd}). For all the datasets except
HET/Malawi and MSM/Beijing, the posterior mean was either outside, or very
close to the border of, the 50\% \gls{HPD}. This indicates that the posterior
distributions were diffuse with heavy tails, rather than having most of their
mass around the mode. When the value $\gls{m} = 1$ was disallowed by the prior, 
the separation between the \gls{IDU} datasets and the others became more
striking (\cref{fig:abchpdm2}). All three \gls{IDU} datasets had estimated
\gls{alpha} values at or above
    \Sexpr{pmean2[c("zetterberg2004", "niculescu2015", "bctree"), round(min(alpha), 2)]}.
The estimate for the MSM/Beijing data was slightly lower
    (\Sexpr{pmean2["wang2015", round(alpha, 2)]}) and
the estimates for the seven remaining non-\gls{IDU} datasets were bounded above by
    \Sexpr{pmean2[c("novitsky2014", "mccormack2002", "grabowski2014", "kao2011", "little2014", "li2015", "cuevas2009"), round(max(alpha), 2)]}.}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{realdata-genes}
    \caption[
        Posterior means and 50\%/95\% \gls{HPD} intervals for parameters of the
        \gls{BA} network model, fitted to three \gls{HIV} datasets where both
        \textit{gag} and \textit{env} genes were sequenced.
    ]{
        Posterior means (points) 50\% \gls{HPD} intervals (notches), and 95\%
        \gls{HPD} intervals (lines) for parameters of the \gls{BA} network model,
        fitted to three \gls{HIV} datasets where both \textit{gag} and
        \textit{env} genes were sequenced.
    }
    \label{fig:genes}
\end{figure}

>>>>

\subsection{Using synthetic data to investigate identifiability and sources of
estimation error}

@@

The accuracy of classifiers for \gls{I} \del{varied significantly
with}\add{appeared to be strongly influenced by} the number of tips in the
tree. For 100-tip trees, the average $R^2$ values were
  \Sexpr{kernel.I[ntip == 100, round(mean(rsquared), 2)]},
  \Sexpr{nltt.I[ntip == 100, round(mean(rsquared), 2)]}, and
  \Sexpr{sackin.I[ntip == 100, round(mean(rsquared), 2)]}
for the tree kernel, \gls{nltt}, and Sackin's index respectively. For 500-tip
trees, the values increased to
  \Sexpr{kernel.I[ntip == 500, round(mean(rsquared), 2)]},
  \Sexpr{nltt.I[ntip == 500, round(mean(rsquared), 2)]}, and
  \Sexpr{sackin.I[ntip == 500, round(mean(rsquared), 2)]}.
Finally, the performance of classifiers for $N$ depended heavily on the
epidemic scenario. The $R^2$ of the \gls{kSVR} classifier ranged from
  \Sexpr{kernel.N[,round(min(rsquared), 2)]}
for the smallest epidemic and smallest sample size, to
  \Sexpr{kernel.N[,round(max(rsquared), 2)]}
for the largest. Likewise, $R^2$ for the \gls{nltt}-based SVR ranged from 
  \Sexpr{nltt.N[,round(min(rsquared), 2)]}
to
  \Sexpr{nltt.N[,round(max(rsquared), 2)]}.
Sackin's index did not accurately classify $N$ in any scenario, with an average
$R^2$ of
  \Sexpr{sackin.N[,round(mean(rsquared), 2)]}
and little variation between scenarios.

@@

\begin{table*}[ht]
    \centering
    \input{\tablepath/abc-hpd.tex}
    \caption[
        Average posterior mean point estimates and 95\% \gls{HPD} interval
        widths for \gls{BA} model parameter estimates obtained with
        \software{netabc} on simulated data.
    ]{
        Average posterior mean point estimates and 95\% \gls{HPD} interval
        widths for \gls{BA} model parameter estimates obtained with
        \software{netabc} on simulated data. Three transmission trees were
        simulated under each combination of the listed parameter values, and
        the parameters were estimated with \gls{ABC} without training.
    }
    \label{tab:abchpd}
\end{table*}

\add{Across all simulations, the median [\gls{IQR}] absolute errors of the parameter 
estimates obtained with \software{netabc} were
    \Sexpr{d[,round(median(alpha_error), 2)]} 
    [\Sexpr{d[,round(quantile(alpha_error, 0.25), 2)]} - 
    \Sexpr{d[,round(quantile(alpha_error, 0.75), 2)]}]
for \gls{alpha},
    \Sexpr{d[,round(median(I_error))]} 
    [\Sexpr{d[,round(quantile(I_error, 0.25))]} - 
    \Sexpr{d[,round(quantile(I_error, 0.75))]}]
for \gls{I},
    \Sexpr{d[,round(median(m_error), 2)]} 
    [\Sexpr{d[,round(quantile(m_error, 0.25), 2)]} - 
    \Sexpr{d[,round(quantile(m_error, 0.75), 2)]}]
for \gls{m}, and
    \Sexpr{d[,round(median(N_error))]} 
    [\Sexpr{d[,round(quantile(N_error, 0.25))]} - 
    \Sexpr{d[,round(quantile(N_error, 0.75))]}]
for \gls{N}. These errors comprised, respectively,
    \Sexpr{d[,round(median(alpha_error) * 100 / 2)]}\%,
    \Sexpr{d[,round(median(I_error) * 100 / 4500)]}\%,
    \Sexpr{d[,round(median(m_error) * 100 / 6)]}\%, and
    \Sexpr{d[,round(median(N_error) * 100 / 14500)]}\%
of the regions of nonzero prior density. For \gls{I} and \gls{N}, relative
errors were
    \Sexpr{d[,round(median(I_error * 100 / true_I))]}\%
    [\Sexpr{d[,round(quantile(I_error * 100 / true_I, 0.25))]} - 
    \Sexpr{d[,round(quantile(I_error * 100 / true_I, 0.75))]}\%]
and
    \Sexpr{d[,round(median(N_error * 100 / true_N))]}\%
    [\Sexpr{d[,round(quantile(N_error * 100 / true_N, 0.25))]} - 
    \Sexpr{d[,round(quantile(N_error * 100 / true_N, 0.75))]}\%].
Average 95\% \gls{HPD} interval widths were
    \Sexpr{d[,round(mean(alpha.upper.95 - alpha.lower.95), 2)]},
    \Sexpr{d[,round(mean(I.upper.95 - I.lower.95))]},
    \Sexpr{d[,round(mean(m.upper.95 - m.lower.95), 2)]}, and
    \Sexpr{d[,format(round(mean(N.upper.95 - N.lower.95)), digits=5)]},
representing 
    \Sexpr{d[,round(mean(alpha.upper.95 - alpha.lower.95) * 100 / 2)]}\%,
    \Sexpr{d[,round(mean(I.upper.95 - I.lower.95) * 100 / 4500)]}\%,
    \Sexpr{d[,round(mean(m.upper.95 - m.lower.95) * 100 / 6)]}\%, and
    \Sexpr{d[,round(mean(N.upper.95 - N.lower.95) * 100 / 14500)]}\%
of the nonzero prior density regions. Point estimates of \gls{I} were upwardly
biased: \gls{I} was overestimated in
    \Sexpr{d[,sum(I > true_I)]}
out of 
    \Sexpr{nrow(d)}
simulations
    (\Sexpr{round(d[,sum(I > true_I)] * 100 / nrow(d))}\%).
The estimates for \gls{m} and \gls{N} were similar across all simulations
(median [\gls{IQR}] point estimates
    \Sexpr{d[,round(median(m), 2)]}
    [\Sexpr{d[,round(quantile(m, 0.25), 2)]} - 
    \Sexpr{d[,round(quantile(m, 0.75), 2)]}]
and 
    \Sexpr{d[,round(median(N))]}
    [\Sexpr{d[,round(quantile(N, 0.25))]} - 
    \Sexpr{d[,round(quantile(N, 0.75))]}])
regardless of the true values of any of the \gls{BA} parameters.}

\add{To analyze the effects of the true parameter values on the
accuracy our estimates of \gls{alpha} and \gls{I}, we fitted one \gls{GLM}
for each of these two parameters, with error rate as the dependent variable and
the true parameter values as independent variables. Since the estimates of
\gls{m} and \gls{N} were roughly equal across all simulations
(\cref{fig:abcptm2,fig:abcptm3,fig:abcptm4}), \glspl{GLM} were not fitted for
these parameters. The estimated coefficients are shown in
\cref{tab:glmalpha,tab:glmI}. The \gls{GLM} were fitted using the inverse link
function. That is, if $p$ is the true value of the parameter and $\hat{p}$ is a
random variable representing our estimate of the parameter, the \gls{GLM}
posits a relationship of the form}
\[
    \madd{\E(|p-\hat{p}|) = (\beta_0 + \beta_\alpha \alpha + \beta_I I + \beta_m m)^{-1}},
\]
\add{where the $\beta$'s are coefficients to be fitted. If the true value
\gls{alpha}, say, is increased by one, the \emph{inverse} of the expected
absolute error will increase by $\beta_\alpha$. If $\beta_\alpha$ is positive,
it means that the absolute error decreases as the true value of \gls{alpha}
increases.}

@@

<<glm_tables, echo=FALSE, warning=FALSE, error=FALSE, results="asis">>=
    source("global.R")
    format.glm <- function (df) {
        df.copy <- copy(df)
        df.copy[Parameter == "alpha", Parameter := "\\alpha"]
        df.copy[Parameter != "(Intercept)", Parameter := paste0("$", Parameter, "$")]
        df.copy[,p := paste0("$", sapply(p, pp, eq=FALSE), "$")]
        df.copy[,Estimate := paste0("$", sapply(Estimate, pp, eq=FALSE, thres=NA), "$")]
        df.copy[,`Standard error` := paste0("$", sapply(`Standard error`, pp, eq=FALSE, thres=NA), "$")]
        setnames(df.copy, "p", "$p$-value")
        df.copy
    }

@@

<<marginals, include=FALSE>>=
    worst.alpha <- d[,mean(alpha_error), by=true_alpha][,true_alpha[which.max(V1)]]
    best.alpha <- d[,mean(alpha_error), by=true_alpha][,true_alpha[which.min(V1)]]
    worst.I <- d[,mean(I_error), by=true_I][,true_I[which.max(V1)]]
    best.I <- d[,mean(I_error), by=true_I][,true_I[which.min(V1)]]
    worst.m <- d[,mean(m_error), by=true_m][,true_m[which.max(V1)]]
    best.m <- d[,mean(m_error), by=true_m][,true_m[which.min(V1)]]

@@

The dispersion of the \gls{ABC} approximation to the posterior also varied
between the parameters (\cref{tab:abchpd}). \gls{HPD} intervals around
\gls{alpha} and \gls{I} were often narrow relative to the region of nonzero
prior density, whereas the intervals for \gls{m} and \gls{N} were more widely
dispersed. \Cref{fig:abclow1,fig:abclow2} \del{shows the distributions for one
simulation.} \add{show one- and two-dimensional marginal distributions for a
simulation with relatively low error. Each parameter was chosen based on its
mean error rate across all simulations. The parameters for this simulation were 
    $\gls{alpha} = \Sexpr{best[,true_alpha]}$,
    $\gls{I} = \Sexpr{best[,true_I]}$,
    $\gls{m} = \Sexpr{best[,true_m]}$, and
    $\gls{N} = \Sexpr{best[,true_N]}$.
\Cref{fig:abchigh1,fig:abchigh2} show the equivalent marginals for a different
simulation with relatively high error rates for each individual parameter. The
parameters were
    $\gls{alpha} = \Sexpr{worst[,true_alpha]}$,
    $\gls{I} = \Sexpr{worst[,true_I]}$,
    $\gls{m} = \Sexpr{worst[,true_m]}$, and
    $\gls{N} = \Sexpr{worst[,true_N]}$.
The two-dimensional marginals indicate some dependence between pairs of
parameters, particularly \gls{I} and \gls{N} which show a diagonally shaped
region of high posterior density.}

@@

\del{We applied \software{netabc} to five published HIV datasets (\cref{tab:data}).
and found substantial heterogeneity among the parameter
estimates.} \add{Posterior mean point estimates and 50\% and 95\% \gls{HPD}
intervals for each parameter are shown in \cref{fig:abchpd}.} \del{and marginal
posterior distributions in
\cref{fig:cuevas,fig:li,fig:niculescu,fig:novitsky,fig:wang}}.
\add{\Cref{fig:abchpdm2} shows point estimates and \gls{HPD} intervals obtained
when the value $\gls{m} = 1$ was disallowed by the prior. Since the results
indicated that $\gls{m} = 1$ was the most credible value for several datasets,
all results discussed henceforth are for the prior $\gls{m} \sim
\DiscreteUniform(1, 5)$ unless otherwise stated.}

\begin{figure}[ht]
    \centering
    \includegraphics{realdata-hpd-bc}
    \caption[
        Posterior means and 50\%/95\% \gls{HPD} intervals for parameters of the
        \gls{BA} network model, fitted to eleven
        \gls{HIV} datasets with \software{netabc}.
    ]{
        Posterior means (points), 50\% \gls{HPD} intervals (notches), and 95\%
        \gls{HPD} intervals (lines) for parameters of the \gls{BA} network
        model, fitted to eleven \gls{HIV} datasets with \software{netabc}.
        Legend labels indicate risk group and country of origin. Abbreviations:
        \acrshort{IDU}, \acrlong{IDU}; \acrshort{MSM}, \acrlong{MSM}; HET,
        heterosexual. Note that posterior means can fall outside of \gls{HPD} 
        intervals if the distribution is diffuse.
    }
    \label{fig:abchpd}
\end{figure}

@@

\add{Posterior mean point estimates for the \acrlong{PA} power \gls{alpha}
were all sub-linear, ranging from 
    \Sexpr{pmean1["zetterberg2004", round(alpha, 2)]}
for the \gls{IDU}/Estonia data to
    \Sexpr{pmean1["cuevas2009", round(alpha, 2)]}
for the mixed/Spain data. When aggregated by risk group, the average estimates
were
    \Sexpr{pmean1[c("zetterberg2004", "niculescu2015", "bctree"), round(mean(alpha), 2)]}
for \gls{IDU},
    \Sexpr{pmean1[c("novitsky2014", "mccormack2002", "grabowski2014"), round(mean(alpha), 2)]}
for primarily heterosexual risk, and
    \Sexpr{pmean1[c("wang2015", "kao2011", "little2014", "li2015"), round(mean(alpha), 2)]}
for \gls{MSM}. These values were obtained with \textit{gag} for the datasets
where both \textit{gag} and \textit{env} were sequenced, but the estimates for
\gls{alpha} did not change appreciably between the two genes (\cref{fig:genes}).
There was a large amount of uncertainty associated with these estimates. 95\%
\gls{HPD} intervals were very wide for most datasets, often encompassing almost
the entire range from 0 to 1 (\cref{fig:abchpd}). For all the datasets except
HET/Malawi and MSM/Beijing, the posterior mean was either outside, or very
close to the border of, the 50\% \gls{HPD} intervals. This indicates that the
posterior distributions were diffuse with heavy tails, rather than having most
of their mass around the mode. When the value $\gls{m} = 1$ was disallowed by
the prior, the separation between the \gls{IDU} datasets and the others became
more striking (\cref{fig:abchpdm2}). All three \gls{IDU} datasets had estimated
\gls{alpha} values at or above
    \Sexpr{pmean2[c("zetterberg2004", "niculescu2015", "bctree"), round(min(alpha), 2)]}.
The estimate for the MSM/Beijing data was slightly lower
    (\Sexpr{pmean2["wang2015", round(alpha, 2)]}) and
the estimates for the seven remaining non-\gls{IDU} datasets were bounded above by
    \Sexpr{pmean2[c("novitsky2014", "mccormack2002", "grabowski2014", "kao2011", "little2014", "li2015", "cuevas2009"), round(max(alpha), 2)]}.}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{realdata-genes}
    \caption[
        Posterior means and 50\%/95\% \gls{HPD} intervals for parameters of the
        \gls{BA} network model, fitted to three \gls{HIV} datasets where both
        \textit{gag} and \textit{env} genes were sequenced.
    ]{
        Posterior means (points), 50\% \gls{HPD} intervals (notches), and 95\%
        \gls{HPD} intervals (lines) for parameters of the \gls{BA} network model,
        fitted to three \gls{HIV} datasets where both \textit{gag} and
        \textit{env} genes were sequenced.
    }
    \label{fig:genes}
\end{figure}
